---
title: "Fitness Activity Analysis"
author: "Jeffrey Uslan"
date: "September 14, 2015"
output: pdf_document
---

#Exectuive Summary



Load in relevant libraries and view data
```{r message=FALSE}
library(ggplot2)
library(tidyr)
library(zoo)
library(dplyr)
```

# Loading and preprocessing the data
```{r}
activity<-tbl_df(read.csv(unzip("./repdata-data-activity.zip")))
```

Checking basic summary.
```{r echo=FALSE}
print(summary(activity))
```

# What is mean total number of steps taken per day?
Calculating the total number of steps taken per day.
```{r echo=FALSE}
sum_steps <- activity %>% group_by(date) %>% summarise(daily_steps=sum(steps))
```

 Making a histogram of the total number of steps taken each day.
```{r echo=FALSE}
ggplot(sum_steps,aes(x=daily_steps))+geom_histogram()+theme_bw()+
  xlab("Daily Steps")+ggtitle("Total Daily Steps")
```
Calculating the mean and median of the total number of steps taken per day


```{r echo=FALSE}
 print(sum_steps %>%   na.omit()  %>% summarise(Mean_Daily_Steps=mean(daily_steps),Median_Daily_Steps=median(daily_steps)) )
```


Average daily activity pattern.
```{r warning=FALSE}
ggplot(activity,aes(x=interval,y=steps))+geom_point()+geom_smooth(aes(group=1))+
  xlab("Interval")+ylab("Steps")+ggtitle("Total Daily Steps")
```
 Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
activity$interval=as.factor(activity$interval)
mean_steps <- activity %>% na.omit() %>% group_by(interval) %>% summarise(interval_steps=mean(steps))
print(mean_steps[which(mean_steps$interval_steps==max(mean_steps$interval_steps)),])
```


# Imputing missing values
 Calculate and report the total number of missing values in the dataset 
```{r}
print(apply(activity,2,function(x){
  sum(is.na(x))
})[1])
```
 

 Create a new dataset that is equal to the original dataset but with the missing data filled in. This method imputes using the averagr for that step.
```{r}
imputed_activity<-apply(activity,1,function(x){
  if (is.na(x[1])){
    ind=which(x[3]==mean_steps$interval)
    x[1]=round(mean_steps$interval_steps[ind])
    
  }
  return(x)
}
)

imputed_activity=as.data.frame(t(imputed_activity))
imputed_activity$steps=as.numeric(as.character(imputed_activity$steps))
```

Checking the summaries of pre and post-imputed datasets as a sanity check.
```{r}
summary(imputed_activity$steps)
summary(activity$steps)
```
 Make a histogram of the total number of steps taken each day and 

```{r}
imputed_sum_steps <- imputed_activity %>% group_by(date) %>% summarise(daily_steps=sum(steps))
ggplot(imputed_sum_steps,aes(x=daily_steps))+geom_histogram()+theme_bw()+
  xlab("Daily Steps")+ggtitle("Imputed Total Daily Steps")
```
 Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? 
```{r echo=FALSE}
print(imputed_sum_steps %>%   na.omit()  %>% summarise(Mean_Daily_Steps_Imputed=mean(daily_steps),Median_Daily_Steps_Imputed=median(daily_steps)) )
print(sum_steps %>%   na.omit()  %>% summarise(Mean_Daily_Steps=mean(daily_steps),Median_Daily_Steps=median(daily_steps)) )

```
 


#Weekends vs Weekdays
 Create a new factor variable in the dataset with two levels – 
 “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
imputed_activity$date=as.POSIXct(activity$date)
week_break <- function(date) {
  day <- weekdays(date)
  if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")){
    return("Weekday")
  }
  else { (day %in% c("Saturday", "Sunday"))
    return("Weekend")
  }
}
imputed_activity$week_break<-sapply(imputed_activity$date,FUN=week_break)
```
A panel plot of the average number of steps taken, averaged across all weekday days or weekend days. 


```{r echo=FALSE}
imputed_sum_steps_weekday <- imputed_activity %>% group_by(date,week_break) %>% summarise(daily_steps=sum(steps))
ggplot(imputed_activity,aes(x=interval,y=steps))+geom_point()+geom_smooth(aes(group=1))+
  xlab("Interval")+ylab("Steps")+ggtitle("Total Daily Steps")+facet_grid(week_break~.)
```
