---
title: "Course Project Part 1"
author: "Jeffrey Uslan"
date: "Monday, July 20, 2015"
output: pdf_document
---
```{r echo==FALSE}
library(ggplot2)
```
In this project I will explroe the central limit theorem. I will do so by comparing 1000 simulations 
of 40 random exponentials variables to the theoretical exponential distribution.
##Setting up simulations

```{r}
set.seed(1)
lambda=.2
n <- 1000
dists=NULL
for (i in 1:n){
  dists=cbind(dists,rexp(40,lambda)  )

}
```

# 1. Show the sample mean and compare it to the theoretical mean of the distribution.
```{r, fig.width = 3, fig.height = 3}
means=apply(dists,2,mean)
cum_means <- cumsum(means)/(1:n)

g <- ggplot(data.frame(x = 1:n, y = cum_means), aes(x = x, y = y))
g <- g + geom_hline(yintercept = 1/lambda) + geom_line(size = 2)
g <- g + labs(x = "Number of obs", y = "Cumulative mean")
g
```
##Theoretical mean
```{r}
1/lambda
```
##Mean of simulation
```{r}
mean(means)
```
# 2. Show how variable the sample is (via variance) and compare it  to the theoretical variance of the distribution.
```{r, fig.width = 3, fig.height = 3}
var=apply(dists,2,var)
cum_var <- cumsum(var)/(1:n)


g <- ggplot(data.frame(x = 1:n, y = cum_var), aes(x = x, y = y))
g <- g + geom_hline(yintercept = (1/lambda)^2) + geom_line(size = 2)
g <- g + labs(x = "Number of obs", y = "Cumulative Varaince")
g
```
##Theoretical variance
```{r}
(1/lambda)^2
```
##Mean of simulation
```{r}
mean(cum_var)
```
# 3. Show that the distribution is approximately normal.
Creating a normal distrubtion using the borrowed parameters
```{r}
normals=rnorm(n,mean=mean(means),sd=sd(means))
```
See how closely the normal ditribution lines up with the simulations
```{r, fig.width = 3, fig.height = 3}
ggplot(data=data.frame(x = 1:n, mean = means,normals=normals))+theme_bw()+
  geom_density(aes(x=means),color="blue")+geom_density(aes(x=normals),color="red")
```
