---
title: "Course Project 2"
author: "Jeffrey Uslan"
date: "Tuesday, July 21, 2015"
output: pdf_document
---

```{r echo=FALSE, results="hide", include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(datasets)
library(reshape)
```
# 1. Load the ToothGrowth data and perform some basic exploratory data analyses 

```{r}
data(ToothGrowth)
toothGrowth=ToothGrowth
toothGrowth=tbl_df(toothGrowth)
toothGrowth$dose=as.factor(toothGrowth$dose)

str(toothGrowth)

ggplot(data=toothGrowth)+theme_bw()+geom_density(aes(x=len,color=dose))+facet_grid(supp~.)
```

# 2. Provide a basic summary of the data.
```{r}
summary(toothGrowth)
table(toothGrowth$supp,toothGrowth$dose)

toothGrowth %>% group_by(supp) %>% summarise(min(len),mean(len),max(len))
toothGrowth %>% group_by(dose) %>% summarise(min(len),mean(len),max(len))
```


# 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 
Gather relevant indices for comparing dose. Collecting Confidence intervals
```{r}
ind_.5=which(toothGrowth$dose==.5)
conf_.5=t.test(x = toothGrowth$len[ind_.5])$conf.int[1:2]
ind_1=which(toothGrowth$dose==1)
conf_1=t.test(x = toothGrowth$len[ind_1])$conf.int[1:2]
ind_2=which(toothGrowth$dose==2)
conf_2=t.test(x = toothGrowth$len[ind_2])$conf.int[1:2]
```

The 95% confidence intervals for doses of .5, 1, and 2 respectively.

```{r}
conf_.5
conf_1
conf_2
```



Gather relevant indices for comparing supp. Collecting Confidence intervals
```{r}
OJ_ind=which(toothGrowth$supp=="OJ")
oj_conf=t.test(x = toothGrowth$len[OJ_ind])$conf.int[1:2]
vc_conf=t.test(x = toothGrowth$len[-OJ_ind])$conf.int[1:2]
```

The 95% confidence intervals for supp of OF and VC respectively.
```{r}
oj_conf
vc_conf

```

A t-test to compare the means of the lengths by supp.
```{r}
OJ_ind=which(toothGrowth$supp=="OJ")
 t.test(x = toothGrowth$len[OJ_ind], y = toothGrowth$len[-OJ_ind], alt = "two.sided", paired = FALSE)
```
The low p-value suggest that we have statistical significant evidence to reject the null hypothesis that the true difference in means is equal to 0 and accept the alternative hypothesis that the true difference is not 0.
