---
title: "Fourier Smoothing"
author: "Jeffrey Uslan"
date: "September 21, 2015"
output: ioslides_presentation
---

## Coursera Reproducible Pitch
### See the Web Application here:
- https://jeffreyuslan.shinyapps.io/Fourier_Smoother

This shiny app is intended to showcase the trend finding and de-noising power of Fourier methods.

Two methods are showcased here:

* The first method removes high frequency modes regardless of amiplitude. 
* The second method removes low amplitude modes regardless of frequency.

The threshold cutoff slider selects the lowest percentile to be omitted.

The code for this application can be found here: ow.ly/SwZrv


## Simulated Data


Creating  simulated time series data:

```{r }
len=1000
x=seq(1:len)
ts       <- seq(1/len,1,1/len)

sin_1=.99*sin(1*2*pi*ts)
sin_2=.9*sin(6*2*pi*ts)
sin_4=.5*sin(26*2*pi*ts)
sin_8=.8*sin(133*2*pi*ts)
sin_16=.7*sin(400*2*pi*ts)
noise=rnorm(len,mean=0,sd=.5)
sim_data=sin_1+sin_2+sin_4+sin_8+sin_16+noise

```
## Simulation Plot
```{r}
plot(sim_data,xlab="Time",ylab="Simulated Data")

```


## Summary of Simulated data
```{r}
summary(sim_data)
```

## Analysis - sample code
```{r, eval=FALSE}
X.hold = fft(sim_data)
cutoff=quantile(Mod(X.hold),Threshold)
cutoff_ind=which(Mod(X.hold)<cutoff)

if (length(cutoff_ind)>1) X.hold[cutoff_ind]=0
X.hold.plottable=Mod(X.hold[0:((length(X.hold)-1)/2)])
X.unf=fft(X.hold,inverse=TRUE)/length(sim_data)
X.unf=as.double(X.unf)
```

